---
title: "Part2_chisquare"
author: "Krista Finlay"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
```

###In Part 2 we will do an introduction to chi-square test for independence in R using lizard tail breakage data recorded at Jornada Basin Long Term Ecological Research site.

####Read in the lizard.csv data and take a look (using View (lizards) in the console) to familiarize yourself with the data structure:

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get the counts of each lizard by species in the data:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
#The most commonly observed lizard is UTST, so just explore that species
```

###Make a subset from lizards data that only contains observations for:

- UTST lizards
- only for sites: IBPE (enclosure grasslands site), RABB (rabbit mesquite site) and EAST (east tarbush site) (maybe different flora landscapre, are the proportion of lizards with broken tails the same or different)
- where the tail condition is recorded as either as W (whole) or B (broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

####Find counts and proportions of lizards with broken or whole tails at each site (and put in a table)
- Use the janitor::tabyl() function

```{r}
utst %>% 
  count(site, tail)
#Table it ok
```

####System for contingency format tables

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

#And you can make a table with both counts and proportions, shows percentage of lizards at each site with broken or whole tails
tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
#can update with kable and kableExtra
```


####Is there a significant association between site and tail breakage, do I have enough evidence to reject the null hypothesis that tail breakage is independent of site

###Convert to a contingency table
Use 'column_to_rowname()' to convert entries in a column to stored rownames

```{r}
lizard_count <- tail_counts %>% 
  column_to_rownames(var = "site")
```

####Now we will run a Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_count)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```




